Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-02-20 appendix.tex
move other part into appendix from Appendix B

>>>>

\section{Inductive Rules for Priority Queues}
\label{sec:inductive rules for priority queues}

Given a sequence $l$ of possibly $\textit{put}$, $\textit{rm}$ and $\textit{rmEmpty}$ operations, we say that $l$ contains unmatched $\textit{rm}$ operations, if the number of $\textit{rm}(d)$ operations is larger than the number of $\textit{put}(d,\_)$ operations for some $d$ in $l$. By data-independence, we can detect the existence of unmatched $\textit{rm}$ operations by automata. A automaton $\mathcal{A}_{\textit{unmRm1}}$ is given in \figurename~\ref{fig:automata for unmatched rm1}, where $c_1=\textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, $c_2=c_1 + \textit{ret}(\textit{rm},b)$, and $c_3 = c_1 + \textit{cal}(\textit{put},b,\textit{anyPri}) + \textit{ret}(\textit{rm},b)$. Here $\textit{anyPri}$ is a specific predicates that is satisfied by any priority. $\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed but has never been inserted.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM1.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{unmRm1}}$}
  \label{fig:automata for unmatched rm1}
\end{figure}

An automaton $\mathcal{A}_{\textit{unmRm2}}$ is given in \figurename~\ref{fig:automata for unmatched rm2}, where $c_1 = \textit{cal}(\textit{put},a,\textit{anyPri}),\textit{cal}(rm),$ $\textit{ret}(\textit{put}),\textit{ret}(\textit{rm},a)$, and $c_1 = c_1 + \textit{ret}(\textit{rm},b)$.  $\mathcal{A}_{\textit{unmRm1}}$ is used to detect the existence of a item that has been removed more than once but has been inserted once.

\begin{figure}[htbp]
  \centering
  \includegraphics[width=0.5 \textwidth]{PIC_AUTO_UNMATCHED_RM2.pdf}
%\vspace{-10pt}
  \caption{Automaton $\mathcal{A}_{\textit{unmRm2}}$}
  \label{fig:automata for unmatched rm2}
\end{figure}

The following lemma states that we can detect the existence of unmatched $\textit{rm}$ by $\mathcal{A}_{\textit{unmRm1}}$ and $\mathcal{A}_{\textit{unmRm2}}$.

\begin{restatable}{lemma}{DetectUnmatchedRmByAutomata}
\label{lemma:detect unmatched rm by automata}

Given a data-independent $\mathcal{I}$, there is some unmatched $\textit{rm}$ in some sequence of $\mathcal{I}$, if and only if $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$.
\end{restatable}

For the $\textit{if}$ direction, assume that $( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm1}} \neq \emptyset) \vee ( \mathcal{I} \cap \mathcal{A}_{\textit{unmRm2}} \neq \emptyset )$. Then, there exists a sequence $l \in \mathcal{I}$ and $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$ or $\mathcal{A}_{\textit{unmRm2}}$. If $l$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$, then $l$ contains at least one $\textit{rm}(b)$ operation and no $\textit{put}(b,\_)$ operation; Else, if $l$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$, then $l$ contains at least two $\textit{rm}(b)$ operations and one $\textit{put}(b,\_)$ operation. In both cases, $l$ contains unmatched $\textit{rm}$ operations.

To prove the $\textit{only if}$ direction assume that there exists a sequence $l \in \mathcal{I}$ that contains unmatched $\textit{rm}$ operations. By data-independence, there exists data-differentiated sequence $l'$ and renaming function $r$, such that $l=r(l')$. We need to prove that in $l'$, one of the following cases holds: (1) $l'$ contains $\textit{rm}(d_1)$ for some $d_1$ and does not contain any $\textit{put}(d_1,\_)$, and (2) $l'$ contains one $\textit{put}(d_1,\_)$ operation and more than one $\textit{rm}(d_1)$ operations. We prove this by contradiction. Assume by contradiction that both cases do not hold. Then, for each value $d$ of $l'$, there is either one $\textit{put}(d,\_)$ and one $\textit{rm}(d)$ in $l'$, or one $\textit{put}(d,\_)$ and no $\textit{rm}(d)$ in $l'$. It is obvious that such $l'$ does not contain unmatched $\textit{rm}$ operation, and renaming a sequence that does not contain unmatched $\textit{rm}$ we can only obtain another sequence that does not contain unmatched $\textit{rm}$. This contradicts that $l=r(l')$ has unmatched $\textit{rm}$.

Let $l''$ be obtained from $ll$ by renaming $d_1$ to $b$ and renaming other value to $a$. If the first case holds, $l''$ is accepted by $\mathcal{A}_{\textit{unmRm1}}$; Otherwise, the second case holds, and $l''$ is accepted by $\mathcal{A}_{\textit{unmRm2}}$. \qed
\end {proof}

Therefore, from now on, we can safely assume that there are no unmatched $\textit{rm}$ operations.

\section{Step-by-Step Linearizability of Priority Queues}
\label{sec:step-by-step linearizability of priority queues}

The following lemma states that $\textit{PQueue}$ is closed under projection. This lemma holds since the predicates appears in inductive rules of $\textit{PQueue}$ are closed under projection.

\begin{restatable}{lemma}{PQisClosedUnderProjection}
\label{lemma:PQ is closed under projection}

$\textit{PQueue}$ is closed under projection.

\end{restatable}

%When the context is clear, we shorten priority-value data-independence as data-independence.
For priority queue, the notion of step-by-step linearizability is modified as follows:

%\vspace{-6pt}
\begin{definition}[step-by-step linearizability for priority queue]\label{def:step-by-step linearizability of priority queue}
A data structure $S=R_1, \ldots, R_n$ is said be to step-by-step linearizable, if for any data-differentiated sequence $e$, if $e$ is linearizable w.r.t. $\textit{MR}_{\textit{pri}}$ ($1 \leq i \leq 4$) with witness $x$, we have:

$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$

\noindent if $e$ is linearizable w.r.t. $\textit{MR}_{\textit{pr5}}$ with witness $x$, we have:

$e \setminus x \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket \Rightarrow e \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr5}} \rrbracket$
\end{definition}

Let $\textit{proj}(s)$ denote the projection of $s$ obtained by erasing all the operations of several items from $s$. Here we assume that each $\textit{rm}(\textit{empty})$ is differentiated from each other, and we can consider each $\textit{rm}(\textit{empty})$ as having a unique item.

\begin{restatable}{lemma}{PQasMultiInMRpriforSequence}
\label{lemma:PQ as multi in MRpri for sequence}
$s \in \textit{PQueue}$, if and only if, $\forall s'$, if $s' \in \textit{proj}(s)$, then $s' \in \textit{MR}_{\textit{pri}}$ for some $0 \leq i \leq 5$.
\end{restatable}

The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.

To prove the $\textit{if}$ direction. The $\textit{if}$ direction holds trivially when $s = \epsilon$. When $s \neq \epsilon$. since $s \in \textit{proj}(s)$, we generate sequence $s_1$ from $s$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $s$ has $\textit{rm}(\textit{empty})$ operations: $s_1$ is generated from $s$ by erasing some $\textit{rm}(\textit{empty})$ operation.

\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: $s_1$ is generated from $s$ by erasing the unmatched $\textit{put}$ operation of items last putted in $S_2$.

\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: $s_1$ is generated from $s$ by erasing this unmatched $\textit{put}$ operation.

\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item which is last putted in $S$.

\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: $s_1$ is generated from $s$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
\end{itemize}

Similarly, we obtain $s_2$ from $s_1$, obtain $s_3$ from $s_2$, $\ldots$, until we obtain some $s_m = \epsilon$. It is obvious that $s_m \in \textit{PQueue}$, and by the five rules of $\textit{PQueue}$, it is easy to see that $s_{\textit{m-1}},\ldots,s_1,s \in \textit{PQueue}$. \qed
\end {proof}

\begin{restatable}{lemma}{PQasMultiInMRpriforHistory}
\label{lemma:PQ as multi in MRpri for history}
Given a data-differentiated history $h$, assume that $\textit{PQueue}$ is step-by-step linearizability. $h \sqsubseteq \textit{PQueue}$, if and only if, $\forall h'$, if $h' \in \textit{proj}(h)$, then $h' \sqsubseteq \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.
\end{restatable}

The \textit{only if} direction is a direct consequence of Lemma \ref{lemma:PQ is closed under projection}.

To prove the $\textit{if}$ direction. Since $h \sqsubseteq \textit{PQueue}$, we can see that $h \sqsubseteq l \in \textit{PQueue}$. Each projection $h'$ of $h$ can be linearizable to some projection $l'$ of $l$. By Lemma \ref{lemma:PQ as multi in MRpri for sequence}, $l' \in \textit{MR}_{\textit{prj}}$ for some $0 \leq j \leq 5$. Since $\textit{last}$ of a sequence is uniquely determined, we can see that $l' \in \textit{MR}_{\textit{pri}}$, where $i = \textit{last}(h')$.

The $\textit{if}$ direction holds trivially when $h = \epsilon$. When $h \neq \epsilon$. since $h \in \textit{proj}(h)$, we generate sequence $h_1$ from $h$ as follows:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $h$ has $\textit{rm}(\textit{empty})$ operations: Then $\textit{last}(h) = R_{\textit{pr5}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr5}}$. $h_1$ is generated from $h$ by erasing some $\textit{rm}(\textit{empty})$ operation.

\item[-] If the maximal priority in $h$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in $S_1$ and unmatched $\textit{put}$ operations of some items in $S_2$: Then $\textit{last}(h) = R_{\textit{pr4}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr4}}$. $h_1$ is generated from $h$ by erasing a unmatched $\textit{put}(x,\_)$, such that $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$ for each $y \in S_2$.

\item[-] If the maximal priority in $s$ is unmatched $\textit{put}$ operations of some items in $S_2$ and $\vert S_2 \vert =1$: Then $\textit{last}(h) = R_{\textit{pr3}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr3}}$. $h_1$ is generated from $h$ by erasing this unmatched $\textit{put}$ operation.

\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert \geq 2$: Then $\textit{last}(h) = R_{\textit{pr2}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr2}}$. $h_1$ is generated from $h$ by erasing $\textit{put}(x,\_)$ and $\textit{rm}(x)$, such that for each $y \in S$, $\textit{put}(x,\_)$ does not happen before $\textit{put}(y,\_)$, and $\textit{rm}(x)$ does not happen before $\textit{rm}(y)$.

\item[-] If the maximal priority in $s$ is of matched $\textit{put}$ and $\textit{rm}$ operations of some items in set $S$ and $\vert S \vert = 1$: Then $\textit{last}(h) = R_{\textit{pr1}}$, and $h \sqsubseteq \textit{MR}_{\textit{pr1}}$. $h_1$ is generated from $h$ by erasing matched $\textit{put}$ and $\textit{rm}$ operations of the item in $S$.
\end{itemize}

Similarly, we obtain $h_2$ from $h_1$, obtain $h_3$ from $h_2$, $\ldots$, until we obtain some $h_m = \epsilon$. Assume that for each $i$, when generating $h_i$ from $h_{\textit{i-1}}$, we erase operations of item $d_i$. It is obvious that $h_m = h_{\textit{m-1}} \setminus d_m \sqsubseteq \textit{PQueue}$, and since $h_{\textit{m-1}} \sqsubseteq \textit{MR}_{\textit{prj}_{\textit{m-1}}}$, where $\textit{last}(h_{\textit{m-1}}) = R_{\textit{prj}_{\textit{m-1}}}$, we can see that $h_{\textit{m-1}} \sqsubseteq \textit{PQueue}$. Similarly, we can prove that $h_{\textit{m-1}},\ldots,h_1,h \sqsubseteq \textit{PQueue}$. This completes the proof of the $\textit{if}$ direction. \qed
\end {proof}

To prove that $\textit{PQueue}$ is step-by-step linearizability, we use five lemmas for five rules of $\textit{PQueue}$.

\begin{restatable}{lemma}{Rpr1isStepByStepLinearizability}
\label{lemma:Rpr1 is step-by-step linearizability}

$R_{\textit{pr1}}$ is step-by-step linearizability.

\end{restatable}

Assume that $l$ matches $R_{\textit{pr1}}$ with witness $x$. Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$. Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.

According to $R_{\textit{pr1}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, and $\textit{pri}_x$ is larger than any priority in $u \cdot v \cdot w$. Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$. Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$. Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$. Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_{L1}$ be the shortest prefix of $l'_L$ that contains all operations of $S_1$, and let $l'_{L2}$ be the sequence such that $l'_L = l'_{L1} \cdot l'_{L2}$. It is obvious that if $\textit{put}(x,\textit{pri}_x)$ happens before an operation $o$ in $h$, then $o$ is in $l'_{L2}$. Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.

Since priority queue is closed under projection (Lemma \ref{lemma:PQ is closed under projection}) and $u \cdot v$ contain matched $\textit{put}$ and $\textit{rm}$ operations, we know that $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$. Since $\textit{pri}_x$ is larger than any priority in $h$, it is easy to see that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$. Since the priority queue is empty after executing $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$, we know that $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R \in \textit{PQueue}$.

It remains to prove that $h \sqsubseteq l''$. To prove this, we define a graph $G$ whose nodes are the operations of $h$ and there is an edge from operation $o_1$ to $o_2$, if one of the following case holds

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $o_1$ happens-before $o_2$ in h,

\item[-] the method event corresponding to $o_1$ in $l''$ is before the one corresponding to $o_2$.
\end{itemize}

Assume there is a cycle in $G$. According the the property of interval order and the fact that the order of $l''$ is total, we know that there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$. Then, we can see

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] If $o_1,o_2 \in l'_{L1}$, or $o_2 \in l'_{L1} \wedge o_1 \in l'_{L2}$, or $o_1,o_2 \in l'_{L2}$, or $o_1,o_2 \in L'_R$, then $l'$ contradicts with happen before relation of $h$.

\item[-] If $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$, then since $\textit{put}(x,\textit{pri}_x)$ happens before $o_2$, we know that $o_2 \in l'_{L2}$, which contradicts that $o_2 \in l'_{L1}$.

\item[-] If $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L1}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L1}$, or $o_1 = \textit{rm}(x) \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 \in L'_R \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, or $o_1 = \textit{rm}(x) \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 \in l'_{L2}$, or $o_1 \in L'_R \wedge o_2 = \textit{rm}(x)$, then $l$ contradicts with happen before relation of $h$.

\item[-] If $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$, then since $o_1$ happens before $\textit{put}(x,\textit{pri}_x)$, we know that $o_1 \in S_1$ and then $o_1 \in l'_{L1}$, which contradicts that $o_1 \in l'_{L2}$.
\end{itemize}

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed

\end {proof}

\begin{restatable}{lemma}{Rpr2isStepByStepLinearizability}
\label{lemma:Rpr2 is step-by-step linearizability}

$R_{\textit{pr2}}$ is step-by-step linearizability.

\end{restatable}

Assume that $l$ matches $R_{\textit{pr2}}$ with witness $x$. Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$. Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.

Since $h' \sqsubseteq l'$, there exists a sequence $h_{l'}$ of call, return and operations, such that (1) the projection of $h_{l'}$ into call and return action is $h$, (2) the projection of $h_{l'}$ into operations is $l'$, and (3) for each matched pair of call action $c$ and return action $r$ of $h$, there exists an operation $o$, such that $a$ before $o$, $o$ before $r$, and $o$ is the operation of $c$ and $r$. Intuitively, $h_{l'}$ is obtained from $h$ by adding linearization points.

According to $R_{\textit{pr2}}$, there exist sequences $u$, $v$, and $w$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v \cdot \textit{rm}(x) \cdot w$, where (1) the $\textit{put}$ operations and $\textit{rm}$ operations in $u \cdot v$ are matched, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v \cdot w$, (3) there exists matched $\textit{put}$ and $\textit{rm}$ operations in $u \cdot v \cdot w$ that has priority $\textit{pri}_x$, (4) $\textit{pri}_x$ is larger than priorities of any unmatched put in $u \cdot v \cdot w$, and (5) all $\textit{put}$ operation with priority $\textit{pri}_x$ is in $u$. Let $D_L$ be the set of items in $u \cdot v$, let $D_R$ be the set of items in $w$. Since $u \cdot v$ contain only matched $\textit{put}$ and $\textit{rm}$ operations, we can see that $D_L \cap D_R = \emptyset$. Let $l'_L = l' \vert_{D_L}$ and $l'_R = l' \vert_{D_R}$.

Let $l'_{\textit{L0}}$ be the shortest prefix of $l'_L$, such that every operation in $l'_{\textit{L0}}$ is before the return action of $\textit{put}(x,\textit{pri}_x)$ in $h_{l'}$. Let $\textit{OP}_x = \{ \textit{put}(d,\textit{pri}_x) \vert \textit{put}(d,\textit{pri}_x)$ is an operation of $h\}$. Let $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t$, where $l_t$ is the projection of $l'$ into operations of $\textit{OP}_x$ that are not in $l'_{\textit{L0}}$. Let $\textit{OP}'_x$ be the set of operations in $l_t$. Let $l'_{\textit{L2}}$ be the projection of $l'_L$ into operations of $l'_L$ that are not in $l'_{\textit{L1}}$. Let sequence $l'' = l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \cdot L'_R$.

As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'_L \in \textit{PQueue}$ and $l'_R \in \textit{PQueue}$. It is obvious that $l'_{\textit{L0}} \in \textit{PQueue}$, and then $l'_{\textit{L1}} = l'_{\textit{L0}} \cdot l_t \in \textit{PQueue}$. Since there is no $\textit{rm}(\textit{Empty})$ in $h$ and $\textit{pri}_x$ is the maximal priority of $h$, $l'_{L1} \cdot l'_{L2} \in \textit{PQueue}$. Since all $\textit{put}$ operations with priority $\textit{pri}$ is in $l'_{L1}$, we know that $l'_{L1} \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_{L2} \cdot \textit{rm}(x) \in \textit{PQueue}$. As in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}, we know that $l'' \in \textit{PQueue}$.

To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}. Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$. Then, we can see

\begin{itemize}
\setlength{\itemsep}{0.5pt}

\item[-] $o_1 \in L'_R \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2} \vee o_2 = \textit{rm}(x))$, or $o_1 = \textit{rm}(x) \wedge (o_2 \in l'_{L1} \vee o_2 = \textit{put}(x,\textit{pri}_x) \vee o_2 \in l'_{L2})$: Then $l$ contradicts with happen before relation of $h$.

\item[-] $o_1, o_2 \in l'_R$ or $o_1, o_2 \in l'_{L2}$: Then $l'$ contradicts with happen before relation of $h$.

\item[-] $o_1, o_2 \in l'_{L1}$: If $(o_1, o_2 \in l'_{L0}) \vee (o_1,o_2 \in \textit{OP}'_x)$, then $l'$ contradicts with happen before relation of $h$.

    Else, it is the case that $o_1 \in \textit{OP}'_x \wedge o_2 \in l'_{L0}$. We can see that $o_1$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_1$ is after the return action of $\textit{put}(x,\textit{pri}_x)$. Since $o_1 <_{\textit{hb}} o_2$, it is easy to see that $\textit{put}(x,\textit{pri}_x) <_{\textit{hb}} o_2$, and then $o_2 \notin l'_{L0}$, which contradicts that $o_2 \in l'_{L0}$.

\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_{L1}$: Since $o_1 <_{\textit{hb}} o_2$, it is obvious that $o_2 \notin l'_{L1}$, contradicts that $o_2 \in l'_{L1}$.

\item[-] $o_1 \in l'_{L2} \wedge o_2 \in l'_{L1}$: If $o_2 \in l'_{L0}$, then $l'$ contradicts with happen before relation of $h$.

    Else, it is the case that $o_2 \in \textit{OP}'_x$. $o_2$ interleaves with $\textit{put}(x,\textit{pri}_x)$ and the linearization point of $o_2$ is after the return action of $\textit{put}(x,\textit{pri}_x)$. Since $o_1 <_{\textit{hb}} o_2$, the return action of $o_1$ is before the return action of $\textit{put}(x,\textit{pri}_x)$. Then it is easy to see that $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.

\item[-] $o_1 \in l'_{L2} \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: Then $o_1 \in l'_{L0}$, contradicts that $o_1 \in l'_{L2}$.
\end{itemize}

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
\end {proof}

\begin{restatable}{lemma}{Rpr3isStepByStepLinearizability}
\label{lemma:Rpr3 is step-by-step linearizability}

$R_{\textit{pr3}}$ is step-by-step linearizability.

\end{restatable}

Assume that $l$ matches $R_{\textit{pr3}}$ with witness $x$. Assume that  Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$. Then, let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.

According to $R_{\textit{pr3}}$, there exist sequences $u$ and $v$, such that $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, where $\textit{pri}_x$ is the maximal priority in $u \cdot v$, and $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$ (may equal priority of some unmatched $\textit{put}$ operations in $u \cdot v$). Let $S_1$ be the set of all operations of $h$ that happen before $\textit{put}(x,\textit{pri}_x)$, let $l'_1$ be the shortest prefix of $l'$ that contains all operations of $S_1$, and let $l'_2$ be the sequence such that $l' = l'_1 \cdot l'_2$. Let sequence $l'' = l'_1 \cdot \textit{put}(x,\textit{pri}_x) \cdot l'_2$.

Since $\textit{pri}_x$ is larger than priority of any matched $\textit{put}$ operations in $u \cdot v$, it is easy to see that $l'' \in \textit{PQueue}$. It remains to prove that $h \sqsubseteq l''$. To prove $h \sqsubseteq l''$, we define graph $G$ as in Lemma \ref{lemma:Rpr1 is step-by-step linearizability}. Assume that there is a cycle in $G$, then there must exists $o_1$ and $o_2$, such that $o_1$ happens-before $o_2$ in $h$, but the corresponding method events are in the opposite order in $l''$. Then, we can see

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $o_1,o_2 \in l'_1$, or $o_1,o_2 \in l'_2$, or $o_1 \in l'_2 \wedge o_2 \in l'_1$: Then $l'$ contradicts with happen before relation of $h$.

\item[-] $o_1 \in l'_2 \wedge o_2 = \textit{put}(x,\textit{pri}_x)$: $o_1 \in S_1$, and then $o_1 \in l'_1$, which contradicts that $o_1 \in l'_2$.

\item[-] $o_1 = \textit{put}(x,\textit{pri}_x) \wedge o_2 \in l'_1$: We can see that in $l'$, $o_2$ must after every operation in $S_1$, and then $o_2 \notin l'_1$, contradicts that $o_2 \in l'_1$.
\end{itemize}

Therefore, we know that $G$ is acyclic, and then we know that $h \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. \qed
\end {proof}

The following lemma states that, from linearization of sub-histories, we can merge them and obtain a linearization (regardless of whether it belongs to sequential specification) of the whole history.

\begin{restatable}{lemma}{MergeTwoLinearization}
\label{lemma:merge two linearization}

Given a history $h$ and operation set $S_1$, $S_2$, let $h_1 = h \vert{S_1}$ and $h_2 = h \vert{S_2}$. Assume that $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, and $S_1 \cup S_2$ contains all operations of $h$. Then, there exists a sequence $l$, such that $h \sqsubseteq l$, $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$.
\end{restatable}

Given a history $h$ and a operation $o \in S_2$, let $\textit{MB}(o) = \{ o' \vert o' \in S_1$ and $o'$ happens before $o$ in $h \}$, let $\textit{SBI}(o) = \textit{min}\{ i \vert l_1[0,i]$ contains all elements of $\textit{MB}(o) \}$.

Let $l = s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[n] \cdot s_{\textit{n+1}}$ be generated as follows, where $n = \vert l_2 \vert$:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $s_1 = l_1[0, \textit{SBI}(l_2[1])]$,

\item[-] If $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ already contains $l_1(\textit{SBI}(l_2[\textit{i+1}]))$, then $s_{\textit{i+1}} = \epsilon$. Otherwise, $s_{\textit{i+1}}$ is a subsequence of $l_1$, which starts from the next of last elements of $s_1 \cdot l_2[1] \cdot \ldots \cdot l_2[i]$ in $l_1$ and ends in $l_1(\textit{SBI}(l_2[\textit{i+1}]))$.
\end{itemize}

It is obvious that $l \vert{S_1} = l_1$ and $l \vert{S_2} = l_2$, and it remains to prove that $h \sqsubseteq l$. We prove this by contradiction. Assume that $h$ is not linearizable with respect to $l$. Then there must be two operations of $h$, such that $o_1 <_{hb} o_2$ in $h$ but $o_2$ before $0_1$ in $l$. Since $l \vert{S_1} = l_1$, $l \vert{S_2} = l_2$, and $h_1 \sqsubseteq l_1$, $h_2 \sqsubseteq l_2$, it is easy to see that it is impossible that $o_1,o_2 \in S_1$ or $o_1,o_2 \in S_2$. There are only two possibilities:

\begin{itemize}
\setlength{\itemsep}{0.5pt}
\item[-] $o_1 \in S_1 \wedge o_2 \in S_2$. Then we can see that $o_2=l_2[i]$ and $o_1 \in s_j$ for some $i < j$. Since $o_1 <_{hb} o_2$, we know that $o_1 \in \textit{SBI}(o_2)$. By the construction of $l$, we know that $o_1$ must be in $s_k$ for some $k \leq i$, contradicts that $o_1 \in s_j$ with $i < j$.

\item[-] $o_1 \in S_2 \wedge o_2 \in S_1$. Then we can see that $o_2 \in s_i$ and $o_1 = l_2[j]$ for some $i \leq j$. It is easy to see that this leads to contradiction when $i = j$. For the case of $i \neq j$, we need to satisfy the following requirements: (1) $o_1$ ($l_2[j]$) does not happen before $l_2[i]$, (2) $l_2[i]$ does not happen before $o_2$, (3) $o_2$ is either overlap or happens before $o' \in \textit{MB}(l_2[i])$, and (4) $o' <_{hb} l_2[i]$. By enumeration we can see that it is impossible that above four conditions be satisfied while $o_1 <_{hb} o_2$.
\end{itemize}

This completes the proof of this lemma. \qed
\end {proof}

\begin{restatable}{lemma}{Rpr4isStepByStepLinearizability}
\label{lemma:Rpr4 is step-by-step linearizability}

$R_{\textit{pr4}}$ is step-by-step linearizability.

\end{restatable}

Assume that $l$ matches $R_{\textit{pr4}}$ with witness $x$. Let $h$ be a data-differentiated history, and $l$ be an operation sequence such that $h \sqsubseteq l$. Let $h'=h \setminus x$ and assume that $h' \sqsubseteq \llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$. Let $l'$ be a sequence in $\llbracket R_{\textit{pr0}},\ldots,R_{\textit{pr4}} \rrbracket$ such that $h' \sqsubseteq l'$.

According to $R_{\textit{pr4}}$, there exist sequences $u$ and $v$, such that (1) $l=u \cdot \textit{put}(x,\textit{pri}_x) \cdot v$, (2) $\textit{pri}_x$ is the maximal priority in $u \cdot v$, (3) $\textit{pri}_x$ equals the maximal priority of matched $\textit{put}$ operations in $u \cdot v$, (4) all matched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$ and (5) all the other unmatched $\textit{put}$ operations with priority $\textit{pri}_x$ is in $u$.
